Designing Types for R, Empirically

### 2020-10-14 paper/main.tex
edits, tables

\AT{Let's discuss alternative strategies for consolidating types here.}

@@

To connect this discussion with actual R types, we chose the \AT{which?} function from the \AT{which?} package, and show what the type of the function would be when using each of these strategies.
We refer the reader to Figure~\ref{tbl:real-function-different-merge-strats}.
\AT{We could use stringr::str\_extract\_all?}

\begin{table}[H]
	\centering
	 \resizebox{.98\linewidth}{!}{
		\begin{tabular}{llrrr}
			\toprule
			Merge Strategy & Function Type \\
			\midrule
			Union of Arrows & \\
			Hybrid &  \\
			Arrow of Unions & \\
			\bottomrule
	\end{tabular}}
	\caption{Types of \AT{package::function} with different type merge strategies. }\label{tbl:real-function-different-merge-strats}
\end{table}

>>>>

@@

\begin{figure}[!h]  \centering
	\includegraphics[width=.8\linewidth]{plots/merge_strategy_comp.pdf}
	\caption{Comparison of top-level function type counts across different merge strategies.}\label{fig:mergecomp}
\end{figure}

% To connect this discussion with actual R types, we chose the \code{discretize} function from the \code{arules} package, and show what the type of the function would be when using each of these strategies.
% The purpose of the function is to convert a continuous variable into a categorical variable, using several unsupervised learning methods.
To connect this discussion with actual R types, we chose the \code{correlation} function from the \code{agricolae} package, and show what the type of the function would be when using each of these strategies.
\code{agricolae::correlation} obtains the coefficients of correlation and p-value between all variables of some input data table, and uses a method of the user's choosing.
It returns the correlation matrix and the probability together in a list.
We refer the reader to Figure~\ref{tbl:real-function-different-merge-strats}.
We see that the hybrid approach produces a much smaller type, with only two arrow types in the top-level union.
Moreover, we see that no real precision is lost when further reducing the function type to a pure arrow of unions, as the two options in the type obtained with the hybrid approach are not very informative.

\begin{table}[H]
	\centering
	 \resizebox{.98\linewidth}{!}{
		\begin{tabular}{llrrr}
			\toprule
			Merge Strategy & Function Type \\
			\midrule
			Union of Arrows & \makecell[tl]{
				\FUN{\VEC{\dbl}, \VEC{\dbl}, \chr, \chr}{\nul} | \\
				\FUN{\VEC{\int}, \VEC{\dbl}, \chr, \chr}{\nul} | \\
				\FUN{\CLASS{{\tt data.frame}}, \nul, \VEC\chr, \chr}{\LIST{\CLASS{{\tt matrix}} \; | \; \dbl}} | \\
				\FUN{\CLASS{{\tt data.frame}}, \nul, \chr, \chr}{\LIST{\CLASS{{\tt matrix}} \; | \; \dbl}} | \\
				\FUN{\VEC\dbl, \CLASS{{\tt data.frame}}, \VEC\chr, \chr}{\LIST{\CLASS{{\tt matrix}} \; | \; \dbl}} | \\
				\FUN{\VEC\dbl, \CLASS{{\tt data.frame}}, \chr, \chr}{\LIST{\CLASS{{\tt matrix}} \; | \; \dbl}}
			} \\
		\addlinespace
			Hybrid & \makecell[tl]{
				\FUN{\VEC{\dbl}, \VEC{\dbl}, {\chr}, {\chr}}{\nul} | \\
				\FUN{\CLASS{{\tt data.frame}} \; | \; \VEC{\dbl}, ? \; \CLASS{{\tt data.frame}}, \VEC{\chr}, {\chr}}{\LIST{\CLASS{{\tt matrix}} \; | \; \dbl}}} \\
			\addlinespace
			Arrow of Unions & \FUN{\CLASS{{\tt data.frame}} \; | \; \VEC{\dbl}, ? \; \VEC{\dbl} \; | \; \CLASS{{\tt data.frame}}, \VEC{\chr}, {\chr}}{\, ? \; \LIST{\CLASS{{\tt matrix}} \; | \; \dbl}} \\
			\bottomrule
	\end{tabular}}
	\caption{Types of \code{agricolae::correlation} with different type merge strategies. }\label{tbl:real-function-different-merge-strats}
\end{table}

On the other hand, there are real cases where a meaningful loss of precision occurs.
Consider instead the \code{rename} function from the \code{dplyr} package, with types shown in Figure~\ref{tbl:real-function-different-merge-strats-2}.
This function takes a data frame and renames selected columns.
We see that with both the union-of-arrows and hybrid approach, the types are the same (with only three unique function signatures observed), and can glean from these types that \code{dplyr::rename} produces a data frame of the {\it same class} as the input.
This information is lost in the arrow-of-unions merge strategy.
That said, once we extend the type system with proper user-defined type hierarchies, the union-of-arrows type will be much more clear.

\begin{table}[H]
	\centering
	\resizebox{.98\linewidth}{!}{
		\begin{tabular}{llrrr}
			\toprule
			Merge Strategy & Function Type \\
			\midrule
			Union of Arrows &  \makecell[tl]{
		      \FUN{\CLASS{{\tt data.frame}}, ...}{\CLASS{{\tt data.frame}}}	| \\
             \FUN{\CLASS{{\tt data.frame, grouped\_df, tbl, tbl\_df}}, ...}{\CLASS{{\tt data.frame, grouped\_df, tbl, tbl\_df}}} | \\
             \FUN{\CLASS{{\tt data.frame, tbl, tbl\_df}}, ...}{\CLASS{{\tt data.frame, tbl, tbl\_df}}}
			} \\
		\addlinespace
			Hybrid & \makecell[tl]{
				\FUN{\CLASS{{\tt data.frame}}, ...}{\CLASS{{\tt data.frame}}}	| \\
				\FUN{\CLASS{{\tt data.frame, grouped\_df, tbl, tbl\_df}}, ...}{\CLASS{{\tt data.frame, grouped\_df, tbl, tbl\_df}}} | \\
				\FUN{\CLASS{{\tt data.frame, tbl, tbl\_df}}, ...}{\CLASS{{\tt data.frame, tbl, tbl\_df}}}
			} \\
		\addlinespace
			Arrow of Unions & \makecell[tl]{
				$\langle \CLASS{{\tt data.frame}} \;|\; \CLASS{{\tt data.frame, tbl, tbl\_df}} \;|\; \CLASS{{\tt data.frame, grouped\_df, tbl, tbl\_df}, ...} \rangle$ $\rightarrow$ \\
				$\CLASS{{\tt data.frame}} \;|\; \CLASS{{\tt data.frame, tbl, tbl\_df}} \;|\; \CLASS{{\tt data.frame, grouped\_df, tbl, tbl\_df}}$
			} \\
			\bottomrule
	\end{tabular}}
	\caption{Types of \code{X::Y} with different type merge strategies. }\label{tbl:real-function-different-merge-strats-2}
\end{table}
